---
title: "01_run"
author: "Sean Maden"
date: "2023-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("..")
setwd("..")
load("./env/05_summaries/01_run_script.RData")
```

```{r}
y <- experimentData$y.table
z <- experimentData$z
ptrue <- experimentData$p.true
```

Summaries of contents of input data types.

# Data types

## Z ref S4

This is the ABIS-seq reference from S4 cohort. It was available in the manuscript supplement.

## Ptrue S13

This is the true cell type proportions from flow cytometry. It was available in the manuscript supplement.

## Y S13

This is the bulk expression from S13 cohort. It was available from GEO GSE107011 as the file "GSE107011_Processed_data_TPM.txt".

# Summaries

## Cell types

Z ref S4

```{r}
data <- z
cellTypes <- colnames(data)
cellTypes
length(unique(cellTypes))
```

Ptrue S13

```{r}
# data <- ptrue[,1]
cellTypes <- ptrue[,1]
cellTypes
length(unique(cellTypes))
```

Y bulk S13

```{r}
data <- z
cellTypes <- colnames(data)
cellTypes
length(unique(cellTypes))
```

## Expression

# Session info

```{r}
sessionInfo()
```