---
title: "R Notebook"
author: Sean Maden
output: html_notebook
---

Notebook to perform A/B testing on Sosina et al bulk expression datasets.

Run this from base path on local: sosina_reproducibility_notes/sm_reproducibility_report/

```{r, messages = F}
libv <- c("here", "SummarizedExperiment", "MuSiC")
sapply(libv, library, character.only = T)
```

Source fig2 data.

```{r, messages = F}
scripts.path <- "scripts"
source(file.path(scripts.path,"chunk2_prepare-predictions.R"))
source(file.path(scripts.path, "chunk2_prepare-cellsize-osm.R"))
source(file.path(scripts.path,"chunk_predict-plot-music-cellsize-osm.R"))
```

Load bulk datasets.

```{r}
base.path <- "count_data"
rse.counts.all.path <- file.path(base.path, "NAc_Nicotine_hg38_rseGene_rawCounts_allSamples_n239.rda")
rse.counts.postqc.path <- file.path(base.path, "NAc_Nicotine_hg38_rseGene_rawCounts_postQCSamples_n223.rda")
rse.rpkm.all.path <- file.path(base.path, "rpkmCounts_NAc_NicotineGrant_hg38_Sept28_2017_n239.rda")
rse.rpkm.all.newpath <- file.path(base.path, "rse-rpkmCounts_NAc_NicotineGrant_hg38_Sept28_2017_n239.rda")
```

Manage RPKM counts.

```{r}
if(!file.exists(rse.rpkm.all.newpath)){
  rpkm.all <- get(load(rse.rpkm.all.path))
  # make rse from rpkm gene counts
  rse.rpkm <- SummarizedExperiment(assays = geneRpkm, rowData = geneMap) %>% as("RangedSummarizedExperiment")
  # save rse from rpkm gene counts
  save(rse.rpkm, file = rse.rpkm.all.newpath)
  # cleanup
  rm(geneRpkm); rm(geneMap); rm(jMap); rm(jRpkm)
  gc()
} else{
  rse.rpkm <- get(load(rse.rpkm.all.newpath))
}
```

Load rse counts objects.

```{r}
rse.all <- get(load(rse.counts.all.path))
rse.qc <- get(load(rse.counts.postqc.path))
```

# Compare bulk datasets

Dataset comparisons

```{r}
ncol(rse.all)
```

```{r}
ncol(rse.qc)
```

```{r}
ncol(rse.rpkm)
```

```{r}
sample.id.all <- "SAMPLE_ID"
sample.id.qc <- "SampleID"
sample.id.rpkm <- "sample.id"
cd.all <- colData(rse.all)
cd.qc <- colData(rse.qc)
cd.rpkm <- data.frame(sample.id = colnames(rse.rpkm))

# filter cd.all on cd.qc
cd.all[,sample.id.all][!cd.all[,sample.id.all] %in% cd.qc[,sample.id.qc]]

# filter cd.rpkm on cd.qc
cd.rpkm[,sample.id.rpkm][!cd.rpkm[,sample.id.rpkm] %in% cd.qc[,sample.id.qc]]

# compare cd.all, cd.rpkm
identical(cd.rpkm[,sample.id.rpkm], cd.all[,sample.id.all])
```

# Plot bulk predictions

Filter bulk markers.

```{r}
rownames(rse.all) <- rowData(rse.all)$Symbol
rownames(rse.qc) <- rowData(rse.qc)$Symbol
rownames(rse.rpkm) <- rowData(rse.rpkm)$Symbol
y.rse.all.25 <- rse.all[rownames(rse.all) %in% rownames(y25),]
y.rse.qc.25 <- rse.qc[rownames(rse.qc) %in% rownames(y25),]
y.rse.rpkm.25 <- rse.rpkm[rownames(rse.rpkm) %in% rownames(y25),]
dim(y.rse.all.25)
dim(y.rse.qc.25)
dim(y.rse.rpkm.25)
```

Get predictions.

```{r}
# predict
# rse.all
music_est_all_unadj <- music_prop(bulk.mtx = assays(y.rse.all.25)[["counts"]], 
                        sc.sce = sc.sce.music, clusters = "Neuron",
                        samples = "Sample", cell_size = NULL)
music_est_all_nrna <- music_prop(bulk.mtx = assays(y.rse.all.25)[["counts"]], 
                        sc.sce = sc.sce.music, clusters = "Neuron",
                        samples = "Sample", cell_size = osm.nrna)
music_est_all_cellarea <- music_prop(bulk.mtx = assays(y.rse.all.25)[["counts"]], 
                        sc.sce = sc.sce.music, 
                        clusters = "Neuron",
                        samples = "Sample",
                        cell_size = osm.cellarea)
# rse.qc
music_est_qc_unadj <- music_prop(bulk.mtx = assays(y.rse.qc.25)[["counts"]], 
                        sc.sce = sc.sce.music, 
                        clusters = "Neuron",
                        samples = "Sample",
                        cell_size = NULL)
music_est_qc_nrna <- music_prop(bulk.mtx = assays(y.rse.qc.25)[["counts"]], 
                        sc.sce = sc.sce.music, 
                        clusters = "Neuron",
                        samples = "Sample",
                        cell_size = osm.nrna)
music_est_qc_cellarea <- music_prop(bulk.mtx = assays(y.rse.qc.25)[["counts"]], 
                        sc.sce = sc.sce.music, 
                        clusters = "Neuron",
                        samples = "Sample",
                        cell_size = osm.cellarea)

# rse.rpkm
music_est_rpkm_unadj <- music_prop(bulk.mtx = assays(y.rse.rpkm.25)[[1]], 
                        sc.sce = sc.sce.music, 
                        clusters = "Neuron",
                        samples = "Sample",
                        cell_size = NULL)
music_est_rpkm_nrna <- music_prop(bulk.mtx = assays(y.rse.rpkm.25)[[1]], 
                        sc.sce = sc.sce.music, 
                        clusters = "Neuron",
                        samples = "Sample",
                        cell_size = osm.nrna)
music_est_rpkm_cellarea <- music_prop(bulk.mtx = assays(y.rse.rpkm.25)[[1]], 
                        sc.sce = sc.sce.music, 
                        clusters = "Neuron",
                        samples = "Sample",
                        cell_size = osm.cellarea)
```

Make new plots.

```{r}
# plot

# rse.all
# get predictions
pred.all.unadj <- music_est_all_unadj$Est.prop.weighted %>% as.data.frame()
pred.all.nrna <- music_est_all_nrna$Est.prop.weighted %>% as.data.frame()
pred.all.cellarea <- music_est_all_cellarea$Est.prop.weighted %>% as.data.frame()
# filter samples
pred.all.unadj <- pred.all.unadj[rownames(pred.all.unadj) %in% prop$samples,]
pred.all.nrna <- pred.all.nrna[rownames(pred.all.nrna) %in% prop$samples,]
pred.all.cellarea <- pred.all.cellarea[rownames(pred.all.cellarea) %in% prop$samples,]
# check sample ids against prop
identical(rownames(pred.all.unadj), prop$samples)
identical(rownames(pred.all.nrna), prop$samples)
identical(rownames(pred.all.cellarea), prop$samples)
# assign new prop variables
prop$music.all.unadj <- pred.all.unadj[,"TRUE"]
prop$music.all.nrna <- pred.all.nrna[,"TRUE"]
prop$music.all.cellarea <- pred.all.cellarea[,"TRUE"]

# rse.qc
# get predictions
pred.qc.unadj <- music_est_qc_unadj$Est.prop.weighted %>% as.data.frame()
pred.qc.nrna <- music_est_qc_nrna$Est.prop.weighted %>% as.data.frame()
pred.qc.cellarea <- music_est_qc_cellarea$Est.prop.weighted %>% as.data.frame()
# check sample ids against prop
identical(rownames(pred.all.unadj), prop$samples)
identical(rownames(pred.qc.unadj), prop$samples)
# assign new prop variables
prop$music.qc.unadj <- pred.qc.unadj[,"TRUE"]
prop$music.qc.nrna <- pred.qc.nrna[,"TRUE"]
prop$music.qc.cellarea <- pred.qc.cellarea[,"TRUE"]

# rse.rpkm
# get predictions
pred.rpkm.unadj <- music_est_all_unadj$Est.prop.weighted %>% as.data.frame()
pred.rpkm.nrna <- music_est_all_nrna$Est.prop.weighted %>% as.data.frame()
pred.rpkm.cellarea <- music_est_all_cellarea$Est.prop.weighted %>% as.data.frame()
# filter samples
pred.rpkm.unadj <- pred.rpkm.unadj[rownames(pred.rpkm.unadj) %in% prop$samples,]
pred.rpkm.nrna <- pred.rpkm.nrna[rownames(pred.rpkm.nrna) %in% prop$samples,]
pred.rpkm.cellarea <- pred.rpkm.cellarea[rownames(pred.rpkm.cellarea) %in% prop$samples,]
# check sample ids against prop
identical(rownames(pred.rpkm.unadj), prop$samples)
identical(rownames(pred.rpkm.nrna), prop$samples)
identical(rownames(pred.rpkm.cellarea), prop$samples)
# assign new prop variables
prop$music.rpkm.unadj <- pred.rpkm.unadj[,"TRUE"]
prop$music.rpkm.nrna <- pred.rpkm.nrna[,"TRUE"]
prop$music.rpkm.cellarea <- pred.rpkm.cellarea[,"TRUE"]

# make plot dataset
dfp <- data.frame(true = prop[,2], 
                  music.qc.unadj = pred.qc.unadj$`TRUE`, 
                  music.qc.nrna = pred.qc.nrna$`TRUE`,
                  music.qc.cellarea = pred.qc.cellarea$`TRUE`,
                  music.all.unadj = pred.all.unadj$`TRUE`, 
                  music.all.nrna = pred.all.nrna$`TRUE`,
                  music.all.cellarea = pred.all.cellarea$`TRUE`,
                  music.rpkm.unadj = pred.rpkm.unadj$`TRUE`,
                  music.rpkm.nrna = pred.rpkm.nrna$`TRUE`,
                  music.rpkm.cellarea = pred.rpkm.cellarea$`TRUE`)

# make plot objects

# rse.qc
ggpt.qc.unadj <- 
  ggplot(dfp, aes(x = music.qc.unadj, y = true)) + 
  geom_point(alpha = 1, col = "springgreen4", shape = 15) +
  geom_abline(slope = 1, intercept = 0) +
  xlim(0, 0.5) + ylim(0, 0.4) + theme_classic()
ggpt.qc.nrna <- 
  ggplot(dfp, aes(x = music.qc.nrna, y = true)) + 
  geom_point(alpha = 1, col = "springgreen4", shape = 15) +
  geom_abline(slope = 1, intercept = 0) +
  xlim(0, 0.5) + ylim(0, 0.4) + theme_classic()
ggpt.qc.cellarea <- 
  ggplot(dfp, aes(x = music.qc.cellarea, y = true)) + 
  geom_point(alpha = 1, col = "springgreen4", shape = 15) +
  geom_abline(slope = 1, intercept = 0) +
  xlim(0, 0.5) + ylim(0, 0.4) + theme_classic()

# rse.all
ggpt.all.unadj <- 
  ggplot(dfp, aes(x = music.all.unadj, y = true)) + 
  geom_point(alpha = 1, col = "springgreen4", shape = 15) +
  geom_abline(slope = 1, intercept = 0) +
  xlim(0, 0.5) + ylim(0, 0.4) + theme_classic()
ggpt.all.nrna <- 
  ggplot(dfp, aes(x = music.all.nrna, y = true)) + 
  geom_point(alpha = 1, col = "springgreen4", shape = 15) +
  geom_abline(slope = 1, intercept = 0) +
  xlim(0, 0.5) + ylim(0, 0.4) + theme_classic()
ggpt.all.cellarea <- 
  ggplot(dfp, aes(x = music.all.cellarea, y = true)) + 
  geom_point(alpha = 1, col = "springgreen4", shape = 15) +
  geom_abline(slope = 1, intercept = 0) +
  xlim(0, 0.5) + ylim(0, 0.4) + theme_classic()

# rse.rpkm
ggpt.rpkm.unadj <- 
  ggplot(dfp, aes(x = music.rpkm.unadj, y = true)) + 
  geom_point(alpha = 1, col = "springgreen4", shape = 15) +
  geom_abline(slope = 1, intercept = 0) +
  xlim(0, 0.5) + ylim(0, 0.4) + theme_classic()
ggpt.rpkm.nrna <- 
  ggplot(dfp, aes(x = music.rpkm.nrna, y = true)) + 
  geom_point(alpha = 1, col = "springgreen4", shape = 15) +
  geom_abline(slope = 1, intercept = 0) +
  xlim(0, 0.5) + ylim(0, 0.4) + theme_classic()
ggpt.rpkm.cellarea <- 
  ggplot(dfp, aes(x = music.rpkm.cellarea, y = true)) + 
  geom_point(alpha = 1, col = "springgreen4", shape = 15) +
  geom_abline(slope = 1, intercept = 0) +
  xlim(0, 0.5) + ylim(0, 0.4) + theme_classic()
```

Check value set identities.

```{r}
identical(dfp$music.qc.unadj, dfp$music.all.unadj)
identical(dfp$music.qc.unadj, dfp$music.rpkm.unadj)
identical(dfp$music.qc.nrna, dfp$music.all.nrna)
identical(dfp$music.qc.nrna, dfp$music.rpkm.nrna)
identical(dfp$music.qc.cellarea, dfp$music.all.cellarea)
identical(dfp$music.qc.cellarea, dfp$music.rpkm.cellarea)
```

Compare plots

```{r}
grid.arrange(ggpt.all.unadj, ggpt.qc.unadj, ggpt.rpkm.unadj, nrow = 1)
```

```{r}
grid.arrange(ggpt.all.nrna, ggpt.qc.nrna, ggpt.rpkm.nrna, nrow = 1)
```

```{r}
grid.arrange(ggpt.all.cellarea, ggpt.qc.cellarea, ggpt.rpkm.cellarea, nrow = 1)
```