---
title: "RNAscope filter proportions, summaries"
author: "Sean Maden"
date: "2023-09-01"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
env.name <- "01_rnascope-filter-proportions-summaries.RData"
env.path <- file.path("outputs", "01_prepare-datasets", env.name)
if(file.exists(env.path)){
  knitr::opts_chunk$set(echo = F, eval = F, include = F)
  save.image(file = env.path)
} else{
  knitr::opts_chunk$set(echo = TRUE)
}
```

```{r}
knitr::opts_chunk$set(echo = TRUE)
libv <- c("here", "nlme", "ggplot2", "gridExtra", "dplyr", 
          "ggforce", "MultiAssayExperiment", "SingleCellExperiment")
sapply(libv, library, character.only = TRUE)
# load rnascope image data
load("~/GitHub/deconvo_method-paper/outputs/01_prepare-datasets/halo-outputs_updated.Rdata")
# rnascope: make sce with image data
sample.id.halo <- "Sample"
img <- halo.output.table
new.img.colnames <- paste0("cell", seq(nrow(img)))
img.data.colnames <- c("Nucleus_Area", "AKT3_Copies", "Cell_Area", 
                       "DAPI_Nucleus_Intensity", "DAPI_Cytoplasm_Intensity")
img.coldata.colnames <- c("SAMPLE_ID", sample.id.halo, "cell_type", "Slide", 
                          "XMin", "XMax", "YMin", "YMax")
img.list <- lapply(img.data.colnames, function(colname){
  new.data <- img[,colname] %>% as.matrix() %>% t()
  colnames(new.data) <- new.img.colnames
  new.data
})
names(img.list) <- img.data.colnames
# gets data as sce
sce.img <- SingleCellExperiment(assays = img.list)
img.coldata <- DataFrame(img[,img.coldata.colnames])
rownames(img.coldata) <- new.img.colnames
colData(sce.img) <- img.coldata
rm(halo.output.table)
# append k label categories
cd <- colData(sce.img)
cd$k2 <- ifelse(grepl("Excit|Inhib", cd$cell_type), "neuron",
                ifelse(grepl("Endo|Oligo|Micro", cd$cell_type), "glial", "NA"))
colData(sce.img) <- cd
# filter na values
rnascope <- sce.img[,!sce.img$k2=="NA"]
gc()

# helper functions
rnascope_properties <- function(sce.img.filter){
  # get filtered rnascope properties in terms of k cell types
  rnascope <- sce.img.filter1
  sample.id.vector <- unique(rnascope$Sample)
  df.rnascope.kdata <- do.call(rbind, lapply("k2", 
                                             function(cell.type.variable){
    do.call(rbind, lapply(sample.id.vector, function(sample.id){
      rnf <- rnascope[,rnascope$Sample==sample.id]
      # proportions
      df.prop <- table(rnf[[cell.type.variable]], rnf$Sample) %>% prop.table()
      # counts
      df.count <- table(rnf[[cell.type.variable]], rnf$Sample)
      # sizes
      df.size <- aggregate(data.frame(area = assays(rnf)[["Nucleus_Area"]][1,]), 
                           by = list(cell_type = rnf[[cell.type.variable]]), FUN = "median")
      # combine and format output
      df.iter <- cbind(df.prop, cbind(df.size, df.count))
      df.iter <- df.iter[,c(1,2,3,5,8)]
      colnames(df.iter) <- c("cell_type", "sample_id", 
                             "true_proportion", "cell_size", "cell_count")
      df.iter$k.label <- cell.type.variable
      df.iter
    }))
  }))
  rownames(df.rnascope.kdata) <- paste0(df.rnascope.kdata$sample_id,";",
                                        df.rnascope.kdata$cell_type, ";",
                                        df.rnascope.kdata$k.label)
  # make transpose
  df.rn.filter <- t(df.rnascope.kdata)
  return(df.rn.filter)
}
```

# Main script

First filter. Nucleus size 78 microns, Circle and Star, 6 cell type labels.

```{r}
max.nucleus.size <- 75
sample.id.vector <- unique(rnascope[[sample.id.halo]])
filter.sce.size <- assays(sce.img)[["Nucleus_Area"]] <= max.nucleus.size
sce.img.filter1 <- sce.img[,filter.sce.size]
df.rn.filter1 <- rnascope_properties(sce.img.filter1)
```

Second filter. Nucleus size 78 microns, Circle and Star, 3 cell type labels

```{r}
cell.types.keep <- c("Excit", "Inhib", "Oligo")
filter.sce.cd <- colData(sce.img.filter1)$cell_type %in% cell.types.keep
sce.img.filter2 <- sce.img.filter1[,filter.sce.cd]
df.rn.filter2 <- rnascope_properties(sce.img.filter2)
```

Third filter. Nucleus size 78 microns, Circle, 2 cell type labels.

```{r}
cell.types.keep <- c("Excit", "Oligo")
filter.sce.cd <- colData(sce.img.filter1)$cell_type %in% cell.types.keep
sce.img.filter3 <- sce.img.filter1[,filter.sce.cd]
df.rn.filter3 <- rnascope_properties(sce.img.filter3)
```

# Summarize cell type proportions by filter

Get tall table of sample ids by filter ids.

```{r, eval = T}
cell.types.vector <- c("glial", "neuron")
list.sce.img <- list("filter1" = sce.img.filter1, 
                     "filter2" = sce.img.filter2, 
                     "filter3" = sce.img.filter3)
df.res.filters <- do.call(rbind, lapply(seq(length(list.sce.img)), function(index){
  filter.name <- names(list.sce.img)[index]
  sce.img <- list.sce.img[[filter.name]]
  sce.img <- sce.img[,sce.img[["k2"]] %in% cell.types.vector]
  sample.id.vector <- unique(sce.img$Sample)
  df.res.all <- do.call(rbind, lapply(sample.id.vector, function(sample.id){
    df.res <- prop.table(table(sce.img[,sce.img$Sample==sample.id]$k2))
    df.res <- as.data.frame(t(as.matrix(df.res)))
    df.res$sample.id <- sample.id
    if(ncol(df.res)==3){df.res}
  }))
  df.res.all$filter.name <- filter.name
  df.res.all
}))
df.res.filters <- as.data.frame(df.res.filters)
```

Scatterplots of glial proportion vs. neuron proportion.

```{r, eval = T}
ggplot(df.res.filters, aes(x = glial, y = neuron, color = sample.id)) + 
  geom_point() + facet_wrap(~filter.name) + xlim(0, 1) + ylim(0, 1) +
  geom_abline(slope = 1, intercept = 0) + ggtitle("Cell proportion") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(df.res.filters, aes(x = glial, y = neuron, color = filter.name)) + 
  geom_point() + facet_wrap(~sample.id) + xlim(0, 1) + ylim(0, 1) +
  geom_abline(slope = 1, intercept = 0) + ggtitle("Cell proportion") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Violin plots of neuron proportion.

```{r, eval = T}
ggplot(df.res.filters, aes(x = sample.id, y = neuron, color = sample.id)) + 
  geom_violin(draw_quantiles = 0.5) + geom_jitter() + facet_wrap(~filter.name) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_hline(yintercept = 0.3)

ggplot(df.res.filters, aes(x = filter.name, y = neuron, color = filter.name)) + 
  geom_violin(draw_quantiles = 0.5) + geom_jitter() + facet_wrap(~sample.id) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_hline(yintercept = 0.3)
```

## Subregion analysis

Get plots with additional facet on the brain subregion.

```{r, eval = T}
df.res.filters$subregion <- gsub(".*_", "", df.res.filters$sample.id)

ggplot(df.res.filters, aes(x = glial, y = neuron, color = sample.id)) + 
  geom_point() + facet_wrap(~subregion*filter.name) + xlim(0, 1) + ylim(0, 1) +
  geom_abline(slope = 1, intercept = 0) + ggtitle("Cell proportion") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(df.res.filters, aes(x = sample.id, y = neuron, color = sample.id)) + 
  geom_violin(draw_quantiles = 0.5) + geom_jitter() + facet_wrap(~filter.name*subregion) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_hline(yintercept = 0.3)

```

# Save environment, session info

```{r}
setwd("..")
setwd("..")
env.name <- "01_rnascope-filter-proportions-summaries.RData"
env.path <- file.path("outputs", "01_prepare-datasets", env.name)
save.image(file = env.path)
#
sessionInfo()
```





