---
title: "01_run"
author: "Sean Maden"
date: "2023-10-17"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
libv <- c("ggplot2", "gridExtra")
sapply(libv, library, character.only = TRUE)
knitr::opts_chunk$set(echo = FALSE)
setwd("..")
setwd("..")
load("./env/08_improvement/01_run_script.RData")
```

Notebook plotting bias differences before and after cell size transformation in 
ABIS-seq immune cell populations.

# Background

## Sample counts

## Flow cytometry data

These were used as the true proportions for cell types.

Flow cytometry proportions table contains "parts" (not fractions or percentages) 
of cell signal amounts (i.e. includes <1 cell not discrete cell counts). mRNA 
yield amounts reflect total cell amounts rather than cell type-specific amounts 
(i.e. one value per sample). 12/13 of the S13 cohort samples have FC data, and 
13/13 have mRNA yields.

Flow cytometry data maps to ABIS-seq Zref cell type identifiers for 17 (of 61 
available) cell types, reflecting all ABIS-seq Zref cell types. Cell types in 
the FCS data (FC proportions table) are not mutually exclusive, but the 
overlapping reference cell type labels are. Plasmablasts and mDCs, which had 
high marker library size/cell size (see notebook `01_abisseq`), have very low 
abundances from flow cytometry.

## Label term definitions

Definitions of cell type labels for predictions.

```{r}
df.def <- data.frame(
  cell.type = c("B.Naive",
                "Basophils.LD",
                "MAIT",
                "mDCs",
                "Monocytes.C",
                "Monocytes.NC.I",
                "Neutrophils.LD",
                "NK",
                "pDCs",
                "Plasmablasts",
                "T.CD4.Naive",
                "T.CD8.Memory",
                "T.CD8.Naive",
                "T.gd.non.Vd2",
                "T.gd.Vd2"),
  description = c("Naive B-cells",
                  "Low-density Basophils",
                  "Mucosal-associated invariant T cells (MAIT)",
                  "Myeloid Dendritic Cells",
                  "Monocytes, type C",
                  "Monocytes, types NC and I combined",
                  "Low-density Neutrophils",
                  "Natural Killer",
                  "Plasmacytoid Dendritic Cells",
                  "Plasmablasts",
                  "CD4+ Naive T-cell",
                  "CD8+ Memory T-cell",
                  "CD8+ Naive T-cell",
                  "T lymphocyte, gamma delta, non-Vdelta2",
                  "T lymphocyte, gamma delta, Vdelta2")
)
knitr::kable(df.def)
```

# Plot

## Cell sizes bar plots

```{r}
dfPlot <- data.frame(cellType = names(cellSizes), size = as.numeric(cellSizes))
dfPlot$cellType <- factor(dfPlot$cellType, levels = dfPlot$cellType[rev(order(dfPlot$size))])
ggplot(dfPlot, aes(x = cellType, y = size)) + geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Scatterplots

Preview plot data

```{r}
knitr::kable(head(dfPlotTallTall))
```

Axis ranges default

```{r}
cellTypesVector <- unique(dfPlotTallTall$cellType)
plotList <- lapply(
  cellTypesVector, function(cellType){
    dfPlotNew <- dfPlotTallTall[dfPlotTallTall$cellType==cellType,]
    ggplot(dfPlotNew, aes(x = trueProportion, y = proportion)) + 
      geom_point() + geom_abline(slope=1,intercept=0) + facet_wrap(~condition) + 
      ggtitle(cellType)
  }
)
names(plotList) <- cellTypesVector
#plotList[[1]]
#plotList[[2]]
#plotList[[3]]
#plotList[[4]]
#plotList[[5]]
#plotList[[6]]
#plotList[[7]]
#plotList[[8]]
#plotList[[9]]
#plotList[[10]]
#plotList[[11]]
#plotList[[12]]
#plotList[[13]]
#plotList[[14]]
#plotList[[15]]
#plotList[[16]]
#plotList[[17]]
plotList[["Plasmablasts"]]
```

Axis ranges 0 to 1

```{r}
cellTypesVector <- unique(dfPlotTallTall$cellType)
plotList <- lapply(
  cellTypesVector, function(cellType){
    dfPlotNew <- dfPlotTallTall[dfPlotTallTall$cellType==cellType,]
    ggplot(dfPlotNew, aes(x = trueProportion, y = proportion)) + 
      geom_point() + geom_abline(slope=1,intercept=0) + facet_wrap(~condition) + 
      ggtitle(cellType) + xlim(0, 1) + ylim(0, 1)
  }
)
names(plotList) <- cellTypesVector
#plotList[[1]]
#plotList[[2]]
#plotList[[3]]
#plotList[[4]]
#plotList[[5]]
#plotList[[6]]
#plotList[[7]]
#plotList[[8]]
#plotList[[9]]
#plotList[[10]]
#plotList[[11]]
#plotList[[12]]
#plotList[[13]]
#plotList[[14]]
#plotList[[15]]
#plotList[[16]]
#plotList[[17]]
plotList[["Plasmablasts"]]
```

Axis ranges 0 to 0.2

```{r}
cellTypesVector <- unique(dfPlotTallTall$cellType)
plotList <- lapply(
  cellTypesVector, function(cellType){
    dfPlotNew <- dfPlotTallTall[dfPlotTallTall$cellType==cellType,]
    ggplot(dfPlotNew, aes(x = trueProportion, y = proportion)) + 
      geom_point() + geom_abline(slope=1,intercept=0) + facet_wrap(~condition) + 
      ggtitle(cellType) + xlim(0, 0.2) + ylim(0, 0.2)
  }
)
names(plotList) <- cellTypesVector
#plotList[[1]]
#plotList[[2]]
#plotList[[3]]
#plotList[[4]]
#plotList[[5]]
#plotList[[6]]
#plotList[[7]]
#plotList[[8]]
#plotList[[9]]
#plotList[[10]]
#plotList[[11]]
#plotList[[12]]
#plotList[[13]]
#plotList[[14]]
#plotList[[15]]
#plotList[[16]]
#plotList[[17]]
plotList[["Plasmablasts"]]
```

## Bias bar plots

Bias difference

```{r}
dfPlot <- dfPlotTallTall
dfPlot$bias.true.pred <- dfPlot$trueProportion-dfPlot$proportion
cellTypesVector <- unique(dfPlotTallTall$cellType)
plotList <- lapply(
  cellTypesVector, function(cellType){
    dfPlotNew <- dfPlot[dfPlot$cellType==cellType,]
    ggplot(dfPlotNew, aes(x = sample.id, y = bias.true.pred)) + 
      geom_bar(stat = "identity") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      ylab("Bias (True - Pred)")
  }
)
names(plotList) <- cellTypesVector
plotList[["Plasmablasts"]]
```

# Session info

```{r}
sessionInfo()
```

