---
title: "02_presentation"
author: "Sean Maden"
date: "2023-11-22"
output: html_document
---

Main presentation notebook showing improvements after cell size transformations.

```{r setup, include=FALSE}
libv <- c("lute", "ggplot2", "gridExtra", "reshape2", "ggforce")
sapply(libv, library, character.only = TRUE)
setwd("..")
setwd("..")
load("./env/08_improvement/01_run_script.RData")
knitr::opts_chunk$set(echo = FALSE)
```

# Background

Cell size scale factors obtained as column sums on Zref (ABIS-seq) cell type marker gene library sizes (column sums).

True proportions (flow cytometry) available for 12 of 13 samples in S13 cohort.

# Analyses

## Cell sizes

Histogram and density plots of cell sizes

```{r}
dfPlotCellSize <- data.frame(cellType = names(cellSizes), size = as.numeric(cellSizes))
dfPlotCellSize$cellType <- factor(dfPlotCellSize$cellType, levels = dfPlotCellSize$cellType[order(dfPlotCellSize$size)])
plot1 <- ggplot(dfPlotCellSize, aes(x = size)) + geom_histogram() + xlab("Cell size") + ylab("Cell type count")
plot2 <- ggplot(dfPlotCellSize, aes(x = size)) + geom_density() + xlab("Cell size") + ylab("Cell amount density")
grid.arrange(plot1, plot2, nrow=2)
```

Stacked histogram of cell sizes

```{r}
plot3 <- ggplot(dfPlotCellSize, aes(x = size, fill = cellType)) + geom_histogram() + xlab("Cell size") + 
  ylab("Cell type count") + guides(fill=guide_legend(ncol=2))
plot3
```

Barplots of cell sizes, with color by cell size.

```{r}
ggplot(dfPlotCellSize, aes(x = cellType, y = size, fill = cellType)) + 
  geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle=45,hjust=1)) + 
  guides(fill=guide_legend(ncol=2))
```

## Proportions

Flow cytometry proportions by sample.

```{r}
dfPlotType <- ptrue |> t() |> melt() |> as.data.frame()
colnames(dfPlotType) <- c("sample.id", "cellType", "proportion.fc")
dfPlotType$is.plasmablast <- ifelse(
  dfPlotType$cellType=="Plasmablasts", "Plasmablasts", "Not-plasmablasts")

plot1 <- ggplot(dfPlotType, aes(x = sample.id, y = proportion.fc, fill = cellType)) + 
  geom_bar(stat = "identity", position = "stack", color = "black") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill=guide_legend(ncol=2))

plot2 <- ggplot(dfPlotType, aes(x = sample.id, y = proportion.fc, fill = is.plasmablast)) + 
  geom_bar(stat = "identity", position = "stack") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill=guide_legend(ncol=2))

plot1
plot2
```

Just plasmablast proportions by sample, with zoom.

```{r}
dfPlotPlasmablasts <- dfPlotType[dfPlotType$cellType=="Plasmablasts",]
dfPlotPlasmablasts$sample.id <- factor(dfPlotPlasmablasts$sample.id,
                                       levels = dfPlotPlasmablasts$sample.id[
                                         order(dfPlotPlasmablasts$proportion.fc)])

plot3 <- ggplot(dfPlotPlasmablasts, 
                aes(x = sample.id, y = proportion.fc)) + 
  geom_bar(stat = "identity", position = "stack") + ggtitle("Plasmablasts") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot3
```

Just non-Plasmablast proportions by sample.

```{r}
dfPlotNotPlasmablasts <- dfPlotPlasmablasts
dfPlotNotPlasmablasts$proportion.fc=1-dfPlotNotPlasmablasts$proportion.fc
dfPlotNotPlasmablasts$cellType <- "Not-plasmablast"

plot1 <- ggplot(dfPlotNotPlasmablasts, aes(x = proportion.fc)) + 
  geom_histogram() + xlab("Proportion non-Plasmablast") + ylab("Sample count")

plot1
```

## Simulation

Setting the conditions. We simulate scRNAseq data from Plasmablasts and 
non-Plasmablasts, at similar proportions to the ABIS FC proportions.

```{r}
setwd("..")
setwd("..")
load("./env/08_improvement/02_simulation_script.RData")
```

```{r}
ptrue <- ptrue
cellTypeLabel <- c(rep("Plasmablast", N1), rep("non-Plasmablast", N2))
zref <- zref
scrnaseq <- scrnaseq
ypseudobulk <- ypseudobulk
```

Heatmap of reference, transformed reference, and pseudobulk.

```{r}
```

Reporting pseudobulk results, at 3 zoom levels.

```{r}
plot1 <- ggplot(dfPseudobulk, aes(x = true.plasmablasts, y = Plasmablasts)) + 
  geom_point() + geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~condition) + xlim(0, 0.15) + ylim(0, 0.15) +
  ylab("Predicted") + xlab("True")
plot2 <- ggplot(dfPseudobulk, aes(x = true.plasmablasts, y = Plasmablasts)) + 
  geom_point() + geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~condition) + xlim(0, 0.5) + ylim(0, 0.5) +
  ylab("Predicted") + xlab("True")
plot3 <- ggplot(dfPseudobulk, aes(x = true.plasmablasts, y = Plasmablasts)) + 
  geom_point() + geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~condition) + xlim(0, 1) + ylim(0, 1) +
  ylab("Predicted") + xlab("True")
plot1
plot2
plot3

# multiplot
plot1 <- plot1 + xlab("") + ylab("") + ggtitle("Zoom1")
plot2 <- plot2 + xlab("") + ylab("") + ggtitle("Zoom2")
plot3 <- plot3 + xlab("") + ylab("") + ggtitle("Zoom3")
grid.arrange(plot1, plot2, plot3, ncol = 1, top = "Plasmablasts", 
             left = "Predicted", bottom = "True")
```

## Main experiments

# Conclusions

# Session info

```{r}
sessionInfo()
```
