---
title: "04_simulations"
author: "Sean Maden"
date: "2023-11-22"
output:
  pdf_document: default
  html_document: default
---

Results from pseudobulk simulations using Plasmablast and Non-plasmablast cell 
proportions from Monaco et al 2019.

```{r setup, include=FALSE}
libv <- c("lute", "ggplot2", "gridExtra", "reshape2", "ggforce")
sapply(libv, library, character.only = TRUE)
setwd("..")
setwd("..")
load("./env/08_improvement/01_run_script.RData")
load("./env/08_improvement/02_simulation_script.RData")
knitr::opts_chunk$set(echo = FALSE)
```

# Background

## Simulations 

This simulation used randomized single-cell expression. Two cell type labels 
were used, either "Plasmablasts" for the larger cell type category, or 
"Non-plasmablasts" for the smaller cell type category.

In addition, the following data were drawn from the experiment: 1. Proportions 
were calculated for each category from flow cytometry proportions; 2. Mean 
expression was calculated from bulk RNA-seq TPM data at ABIS-seq marker genes.

## Cell size calculations

Cell size scale factors obtained as the library sizes (column sums) from the 
RNA-seq reference in Monaco et al 2019 (Table S5). Binarized cell sizes 
(Plasmablasts, Non-plasmablasts) for Non-plasmablasts were taken as the median 
of cell sizes.

# Scatter plots of simulation outcomes

Reporting pseudobulk results, at 3 zoom levels.

```{r}
dfPlot <- dfPseudobulkA
plot1 <- ggplot(dfPseudobulkA, aes(x = true.plasmablasts, y = Plasmablasts)) + 
  geom_point(size = 4, alpha = 0.5) + geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~condition) + xlim(0, 0.15) + ylim(0, 0.15) +
  ylab("Predicted") + xlab("True")
plot2 <- ggplot(dfPseudobulkA, aes(x = true.plasmablasts, y = Plasmablasts)) + 
  geom_point(size = 4, alpha = 0.5) + geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~condition) + xlim(0, 0.5) + ylim(0, 0.5) +
  ylab("Predicted") + xlab("True")
plot3 <- ggplot(dfPseudobulkA, aes(x = true.plasmablasts, y = Plasmablasts)) + 
  geom_point(size = 4, alpha = 0.5) + geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~condition) + xlim(0, 1) + ylim(0, 1) +
  ylab("Predicted") + xlab("True")
plot1
plot2
plot3
```
```{r, fig.height = 15, fig.width = 9}
# multiplot
textSize <- 30
plot1 <- 
  plot1 + xlab("") + ylab("") + ggtitle("Zoom1") + 
  theme(text = element_text(size=textSize),
        axis.text.x = element_text(angle=45,hjust=1))
plot2 <- 
  plot2 + xlab("") + ylab("Predicted") + ggtitle("Zoom2") + 
  theme(text = element_text(size=textSize),
        axis.text.x = element_text(angle=45,hjust=1))
plot3 <- 
  plot3 + xlab("") + ylab("True") + ggtitle("Zoom3") + 
  theme(text = element_text(size=textSize),
        axis.text.x = element_text(angle=45,hjust=1))
grid.arrange(plot1, plot2, plot3, ncol = 1, top = "Plasmablasts")
```

# Session info

```{r}
sessionInfo()
```