---
title: "RNAscope pseudobulk shuffle final iteration, cohort1"
author: "Sean Maden"
date: "2023-09-03"
output:
  pdf_document: default
  html_document: default
---

This notebook shows a pseudobulk shuffle experiment in which $S_{deconvolution}$ is shuffled.

# Variables

Descriptions of the variables used.

* `s.data.frame` : is the sample-wise cell sizes table.

* `dfp.tall` : is the output of the experiment function `get_ypb_experiment_series_shuffle()`.

# Load

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
libv <- c("snow", "ggplot2", "dplyr", "parallel", 
          "doParallel", "lute", "dplyr", "scuttle")
sapply(libv, library, character.only = T)
###
point.color <- "goldenrod1"
###
setwd("..")
setwd("..")
env.name <- "02_figde_script.RData"
env.path <- file.path("env", "03_shuffle", env.name)
load(env.path)
```

# High neuron proportion sample

Save formatted plot.

```{r}
new.plot.low <- ggplot(dfp.tall.high, 
       aes(x = neuron.true, y = neuron.pred, shape = matched.id)) + 
  geom_point(size = 5, alpha = 0.55, colour = point.color) +
  geom_abline(slope = 1, intercept = 0) + facet_wrap(~type) + theme_bw() +
  xlab("Known") + ylab("Predicted") +
  ggtitle(paste0("S deconvo id: ", s.scale.sample.high)) + xlim(0.2, 1) + ylim(0.2, 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
new.plot.low
```

```{r, include = FALSE}
setwd("..")
setwd("..")

new.plot.path <- file.path("figures", "03_shuffle", "fig3d.jpg")

jpeg(new.plot.path, width = 5.5, height = 3, units = "in", res = 400)

new.plot.low

dev.off()
```

# Low neuron proportion sample

Save formatted plot. 

```{r}
new.plot.high <- ggplot(dfp.tall.low, aes(x = neuron.true, y = neuron.pred, shape = matched.id)) + 
  geom_point(size = 5, alpha = 0.55, colour = point.color) + 
  geom_abline(slope = 1, intercept = 0) + facet_wrap(~type) + theme_bw() +
  xlab("Known") + ylab("Predicted") + theme(legend.position = "none") +
  ggtitle(paste0("S deconvo id: ", s.scale.sample.id.low)) + xlim(0.2, 1) + ylim(0.2, 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
new.plot.high
```

```{r, include = FALSE}
setwd("..")
setwd("..")

new.plot.filename <- "fig3e.jpg"
new.plot.path <- file.path("figures", "03_shuffle", new.plot.filename)

jpeg(new.plot.path, width = 5, height = 3, units = "in", res = 400)

new.plot.high

dev.off()
```

# Write table

```{r}
# get write table
dfp.tall.write <- dfp.tall
colnames(dfp.tall.write)[5] <- "sample.id"
colnames(dfp.tall.write)[7] <- "index.sample.id"

dfp.tall.write$neuron.abs.error <- abs(dfp.tall.write$neuron.pred-dfp.tall.write$neuron.true)
dfp.tall.write$glial.abs.error <- abs(dfp.tall.write$glial.pred-dfp.tall.write$glial.true)
dfp.tall.write$shuffle.term <- "S_deconvolution"

colnames.write <- c("glial.true", "neuron.true", 
                    "glial.pred", "neuron.pred", 
                    "type", "sample.id", 
                    "glial.abs.error", "neuron.abs.error",
                    "index.sample.id", "shuffle.term")

dfp.tall.write <- dfp.tall.write[,colnames.write]
```

```{r}
setwd("..")
setwd("..")
write.csv(dfp.tall.write, row.names = F, file = "./outputs/03_shuffle/table_s3_sdeconvo.csv")
```

# Write summary statistics

```{r}

```

# Session info

```{r, include = FALSE}
sessionInfo()
```