---
title: "02_rnascope_image-vs-proportion-filter"
author: "Sean Maden"
date: "2023-09-06"
output: html_document
---

```{r setup, include=FALSE}
#

# check for existing env, otherwise run script

#

notebook.path <- getwd()
setwd("..")
inst.path <- getwd()
setwd("..")
project.path <- getwd()
setwd("..")
data.path <- "data"
output.path <- "path"

source("deconvo_method-paper_cohort1_paths")

env.name <- "01_rnascope-filter-proportions-summaries.RData"
env.path <- file.path("outputs", "01_prepare-datasets", env.name)
if(file.exists(env.path)){
  knitr::opts_chunk$set(echo = F, eval = F, include = F)
  save.image(file = env.path)
} else{
  knitr::opts_chunk$set(echo = TRUE)
}
```

```{r}
knitr::opts_chunk$set(echo = TRUE)
libv <- c("here", "nlme", "ggplot2", "gridExtra", "dplyr", 
          "ggforce", "MultiAssayExperiment", "SingleCellExperiment")
sapply(libv, library, character.only = TRUE)
```

Get HALO image data as `singleCellExperiment`.

```{r}
# load rnascope image data
load("~/GitHub/deconvo_method-paper/outputs/01_prepare-datasets/halo-outputs_updated.Rdata")
# rnascope: make sce with image data
sample.id.halo <- "Sample"
img <- halo.output.table
new.img.colnames <- paste0("cell", seq(nrow(img)))
img.data.colnames <- c("Nucleus_Area", "AKT3_Copies", "Cell_Area", 
                       "DAPI_Nucleus_Intensity", "DAPI_Cytoplasm_Intensity")
img.coldata.colnames <- c("SAMPLE_ID", sample.id.halo, "cell_type", "Slide", 
                          "XMin", "XMax", "YMin", "YMax")
img.list <- lapply(img.data.colnames, function(colname){
  new.data <- img[,colname] %>% as.matrix() %>% t()
  colnames(new.data) <- new.img.colnames
  new.data
})
names(img.list) <- img.data.colnames
# gets data as sce
sce.img <- SingleCellExperiment(assays = img.list)
cd.colnames <- c("SAMPLE_ID", "Sample", "BrNum", "Position", "cell_type", "AKT3_Copies", "XMin", "XMax", "YMin", "YMax")
cd.img <- DataFrame(halo.output.table[,cd.colnames])
rownames(cd.img) <- new.img.colnames
colData(sce.img) <- cd.img
```

# Compare output image vs. proportions labels

```{r}
dim(sce.img)
```

```{r}
sample.id <- "Br6432_ant"
experiment.id <- "STAR"
filter.sce <- sce.img$Sample==sample.id & 
  grepl(paste0(".*_", experiment.id, "$"), sce.img$SAMPLE_ID)
sce.filt <- sce.img[,filter.sce]
dim(sce.filt)
```

Plot single image.

```{r}
dfp <- data.frame(nucleus_area = assays(sce.filt)[["Nucleus_Area"]][1,],
                  xmin = sce.filt$XMin, xmax = sce.filt$XMax,
                  ymin = sce.filt$YMin, ymax = sce.filt$YMax,
                  cell_type = sce.filt$cell_type,
                  point.index.id = colnames(sce.filt))
# apply colors
dfp$is.neuron <- ifelse(dfp$cell_type %in% c("Inhib", "Excit"), TRUE, FALSE)
dfp$is.glial <- ifelse(dfp$cell_type %in% c("Oligo", "Astro", "Micro"), TRUE, FALSE)
dfp$is.oligo <- ifelse(dfp$cell_type %in% c("Oligo"), TRUE, FALSE)
dfp$is.other <- ifelse(dfp$cell_type %in% c("Other"), TRUE, FALSE)
```

Show full tissue plots.

```{r}
ggplot(dfp, (aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax))) + 
  geom_rect(alpha = 0.5) + theme_bw()

ggplot(dfp, (aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, color = is.neuron))) + 
  geom_rect(alpha = 0.5) + theme_bw()

ggplot(dfp, (aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, color = is.glial))) + 
  geom_rect(alpha = 0.5) + theme_bw()

ggplot(dfp, (aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, color = is.oligo))) + 
  geom_rect(alpha = 0.5) + theme_bw()

ggplot(dfp, (aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, color = is.other))) + 
  geom_rect(alpha = 0.5) + theme_bw()
```

Show full plots with exclusions.

```{r}
# Helper functions

exclude_coordinates <- function(dfp, exclusion.decile = 1){
  xrange <- max(dfp$xmax)-min(dfp$xmin)
  yrange <- max(dfp$ymax)-min(dfp$ymin)
  xrange.q <- quantile(seq(0, xrange, 1), 
                       probs = seq(0, 1, 0.1))[[1+exclusion.decile]]
  yrange.q <- quantile(seq(0, yrange, 1),
                       probs = seq(0, 1, 0.1))[[1+exclusion.decile]]
  xint.low.q <- min(dfp$xmin)+xrange.q
  xint.high.q <- max(dfp$xmax)-xrange.q
  yint.low.q <- min(dfp$ymin)+yrange.q
  yint.high.q <- max(dfp$ymax)-yrange.q
  # get filtered dfp
  filter.xcoord <- dfp$xmin > xint.low.q & dfp$xmax < xint.high.q
  filter.ycoord <- dfp$ymin > yint.low.q & dfp$ymax < yint.high.q
  dfp.filter <- dfp[filter.xcoord & filter.ycoord,]
  # draw full plot with axis intercept lines indicating exclusions
  new.plot1 <- ggplot(dfp, 
                     (aes
                      (xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax))) + 
    geom_rect(alpha = 0.5) + theme_bw() +
    geom_hline(yintercept = yint.low.q) + geom_hline(yintercept = yint.high.q) +
    geom_vline(xintercept = xint.low.q) + geom_vline(xintercept = xint.high.q) +
    ggtitle(paste0("Quantile = ", 10*exclusion.decile, "%"))
  # highlight excluded points
  dfp$excluded <- ifelse(
    dfp$point.index.id %in% dfp.filter$point.index.id, FALSE, TRUE)
  new.plot2 <- ggplot(dfp, 
                     (aes
                      (xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                        color = excluded))) + 
    geom_rect(alpha = 0.5) + theme_bw() +
    geom_hline(yintercept = yint.low.q) + geom_hline(yintercept = yint.high.q) +
    geom_vline(xintercept = xint.low.q) + geom_vline(xintercept = xint.high.q) +
    ggtitle(paste0("Quantile = ", 10*exclusion.decile, "%"))
  # return
  lr <- list(new.plot1 = new.plot1, 
              new.plot2 = new.plot2, 
              dfp.filter = dfp.filter)
  return(lr)
}

cell_type_barplots <- function(dfp){
  dfp.bp <- as.data.frame(table(dfp$cell_type))
  colnames(dfp.bp) <- c("cell_type", "cells")
  dfp.bp[,2] <- as.numeric(dfp.bp[,2])
  dfp.bp$percent <- paste0(round(100*dfp.bp[,2]/sum(dfp.bp[,2]), 1), "%")
  dfp.bp$sample <- rep("sample", nrow(dfp.bp))
  new.plot1 <- ggplot(dfp.bp, aes(x = sample, y = cells, fill = cell_type)) + 
    geom_bar(stat = "identity", position = "stacked") + 
    theme_bw()
  new.plot2 <- ggplot(dfp.bp, aes(x = cell_type, y = cells, fill = cell_type)) + 
    geom_bar(stat = "identity", position = "dodge") + 
    geom_text(aes(label = percent), vjust = 0) +
    theme_bw()
  lr <- list(new.plot1 = new.plot1, new.plot2 = new.plot2)
  return(lr)
}

```

Show decile exclusions.

```{r}
exclude_coordinates(dfp, 1)$new.plot2
exclude_coordinates(dfp, 2)$new.plot2
exclude_coordinates(dfp, 3)$new.plot2
exclude_coordinates(dfp, 4)$new.plot2
```

Get cell type proportions by decile exclusion.

```{r}
exclude_coordinates(dfp, 1)$new.plot2
exclude_coordinates(dfp, 2)$new.plot2
exclude_coordinates(dfp, 3)$new.plot2
exclude_coordinates(dfp, 4)$new.plot2
```

Show full plots with exclusions and cell type proportions.

```{r}
list.results <- lapply(c(1,2,4), function(decile.filter){
  list.filter.result <- exclude_coordinates(dfp, decile.filter)
  barplots.filter <- cell_type_barplots(list.filter.result$dfp.filter)
  new.filter.plot <- list.filter.result$new.plot2
  new.barplot1 <- barplots.filter$new.plot1 + ggtitle("")
  new.barplot2 <- barplots.filter$new.plot2 + ggtitle("")
  list(new.filter.plot, new.barplot1, new.barplot2)
})
```

Show 3 plots side-by-side.

```{r}
eval.str <- paste0(
  "list.results[[",seq(length(list.results)), "]][[1]],",
  "list.results[[",seq(length(list.results)), "]][[2]],",
  "list.results[[",seq(length(list.results)), "]][[3]]", 
  collapse = ",")
eval.str <- paste0("grid.arrange(", eval.str, ", ncol = length(list.results[[1]]))")
eval(parse(text = eval.str))
```

Show stacked barplots, percentage and cell counts scales.

```{r}
```






