---
title: "05_rnascope_confidence"
author: "Sean Maden"
date: "2023-10-10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
libv <- c("ggplot2")
sapply(libv, library, character.only = TRUE)
knitr::opts_chunk$set(echo = TRUE)
setwd("..")
setwd("..")
load("./outputs/01_mae/sample_qc_df.rda")
```

Shows snRNAseq neuron and glial proportions versus RNAscope confidence annotation filter.
The annotation filter criteria was either Circle or Star combo having a "Low" confidence status (i.e. label status TRUE) versus otherwise (i.e. label status FALSE).

# Summary

Sample ids removed

```{r}
cd.id[cd.id$remove.low==TRUE,]$sample.id
```

Sample ids included

```{r}
cd.id[cd.id$remove.low==FALSE,]$sample.id
```

Neuron proportion among samples removed

```{r}
summary(cd.id[cd.id$remove.low==TRUE,]$proportion.sn.neuron)
```

Neuron proportion among samples included

```{r}
summary(cd.id[cd.id$remove.low==FALSE,]$proportion.sn.neuron)
```

# Plot proportion neuron

```{r}
newplot <- ggplot(cd.id, 
                  aes(x = sample.id, y = proportion.sn.neuron)) + 
  geom_boxplot() + geom_jitter() +
  facet_wrap(~remove.low) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

newplot
```

```{r}
newplot <- ggplot(cd.id, 
                  aes(x = brain.region, y = proportion.sn.neuron)) + 
  geom_boxplot() + geom_jitter() +
  facet_wrap(~remove.low) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

newplot
```

# Scatterplots of glial and neuron proportions

Unlabeled

```{r}
newplot <- ggplot(cd.id, 
                  aes(x = proportion.sn.neuron, y = proportion.sn.glial)) + 
  geom_point(size = 2, alpha = 0.5) + geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~remove.low)

newplot
```

## Label on sample id

```{r}
newplot <- ggplot(cd.id, 
                  aes(x = proportion.sn.neuron, y = proportion.sn.glial,
                      color = sample.id)) + 
  geom_point(size = 2, alpha = 0.5) + geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~remove.low)

newplot
```

```{r}
newplot <- ggplot(cd.id, 
                  aes(x = proportion.sn.neuron, y = proportion.sn.glial,
                      color = sample.id)) + 
  geom_point(size = 2, alpha = 0.5) + geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~remove.low) +
  geom_label(aes(label = sample.id))

newplot
```

## Label on region

```{r}
newplot <- ggplot(cd.id, 
                  aes(x = proportion.sn.neuron, y = proportion.sn.glial,
                      color = brain.region)) + 
  geom_point(size = 2, alpha = 0.5) + geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~remove.low)

newplot
```

```{r}
newplot <- ggplot(cd.id, 
                  aes(x = proportion.sn.neuron, y = proportion.sn.glial,
                      color = brain.region)) + 
  geom_point(size = 2, alpha = 0.5) + geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~remove.low) +
  geom_label(aes(label = brain.region))

newplot
```

# Session info

```{r}
sessionInfo()
```