---
title: "01_runRmseTable"
author: "Sean Maden"
date: "2024-02-12"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("..")
setwd("..")
source("./src/rmse.R")
```

# Parse proportions outputs

Cohort 1 pseudobulk, K2 and K3

```{r}
# proportions cohort 1
setwd("..")
setwd("..")
setwd("..")
# k2 results
load(file.path("./huuki-meyers-et-al-2023_dlpfc-cohort1/env/02_pseudobulk/01_k2.RData"))
proportionsCohort1K2 <- dfp.tall
rm(dfp.tall)
# k3 results
load(file.path("./huuki-meyers-et-al-2023_dlpfc-cohort1/env/02_pseudobulk/02_k3.RData"))
proportionsCohort1K3 <- dfp.tall
rm(dfp.tall)
```

```{r}
# k2 results
cellTypesVector <- c("glial", "neuron")
proportionsCohort1K2 <- proportionsCohort1K2[,c(1:6)]
colnames(proportionsCohort1K2) <- c(
  paste0(cellTypesVector, "_known"),
  paste0(cellTypesVector, "_predicted"),
  c("condition.id", "sample.id")
)
# k3 results
cellTypesVector <- c("glial", "Excit", "Inhib")
proportionsCohort1K3 <- proportionsCohort1K3[,c(1:8)]
colnames(proportionsCohort1K3) <- c(
  paste0(cellTypesVector, "_known"),
  paste0(cellTypesVector, "_predicted"),
  c("condition.id", "sample.id")
)
```

Cohort 2 pseudobulk, K2 and K3

```{r}
# proportions cohort 2
setwd("..")
setwd("..")
setwd("..")
# k2 results
load(file.path("./tran-et-al-2021_dlpfc-cohort2/env/01_pseudobulk/01_k2_mrb_script.RData"))
proportionsCohort2K2 <- dfp.tall
rm(dfp.tall)
# k3 results
load(file.path("./tran-et-al-2021_dlpfc-cohort2/env/01_pseudobulk/01_k3_mrb_script.RData"))
proportionsCohort2K3 <- dfp.tall
rm(dfp.tall)
```

```{r}
# k2 results
cellTypesVector <- c("glial", "neuron")
proportionsCohort2K2 <- proportionsCohort2K2[,c(1:6)]
colnames(proportionsCohort2K2) <- c(
  paste0(cellTypesVector, "_known"),
  paste0(cellTypesVector, "_predicted"),
  c("condition.id", "sample.id")
)
# k3 results
cellTypesVector <- c("glial", "Excit", "Inhib")
proportionsCohort2K3 <- proportionsCohort2K3[,c(1:8)]
colnames(proportionsCohort2K3) <- c(
  paste0(cellTypesVector, "_known"),
  paste0(cellTypesVector, "_predicted"),
  c("condition.id", "sample.id")
)
```

Cohort 1 bulk

```{r}
setwd("..")
setwd("..")
setwd("..")
load(file.path("./huuki-meyers-et-al-2023_dlpfc-cohort1/env/07_adjustment/03_run_adjustment_realbulk_all_script.RData"))

# nnls
proportionsBulk <-
  data.frame(
    glial_known = 1-dfp.wide$neuron.true,
    neuron_known = dfp.wide$neuron.true,
    glial__predicted = 1-dfp.wide$neuron.nnls.scale,
    neuron_predicted = dfp.wide$neuron.nnls.scale,
    sample.id = rownames(dfp.wide),
    condition.id = rep("scale", nrow(dfp.wide)),
    replicate.id = rep("nnls", nrow(dfp.wide))
  )
proportionsBulk <- rbind(proportionsBulk,
                         data.frame(
    glial_known = 1-dfp.wide$neuron.true,
    neuron_known = dfp.wide$neuron.true,
    glial__predicted = 1-dfp.wide$neuron.nnls.noscale,
    neuron_predicted = dfp.wide$neuron.nnls.noscale,
    sample.id = rownames(dfp.wide),
    condition.id = rep("noscale", nrow(dfp.wide)),
    replicate.id = rep("nnls", nrow(dfp.wide))
  ))

# music
proportionsBulk <- rbind(proportionsBulk,
                         data.frame(
    glial_known = 1-dfp.wide$neuron.true,
    neuron_known = dfp.wide$neuron.true,
    glial__predicted = 1-dfp.wide$neuron.music.scale,
    neuron_predicted = dfp.wide$neuron.music.scale,
    sample.id = rownames(dfp.wide),
    condition.id = rep("scale", nrow(dfp.wide)),
    replicate.id = rep("music", nrow(dfp.wide))
  ))
proportionsBulk <- rbind(proportionsBulk,
                         data.frame(
    glial_known = 1-dfp.wide$neuron.true,
    neuron_known = dfp.wide$neuron.true,
    glial__predicted = 1-dfp.wide$neuron.music.noscale,
    neuron_predicted = dfp.wide$neuron.music.noscale,
    sample.id = rownames(dfp.wide),
    condition.id = rep("no", nrow(dfp.wide)),
    replicate.id = rep("music", nrow(dfp.wide))
  ))
# bisque
proportionsBulk <- rbind(proportionsBulk,
                         data.frame(
    glial_known = 1-dfp.wide$neuron.true,
    neuron_known = dfp.wide$neuron.true,
    glial__predicted = 1-dfp.wide$neuron.bisque.scale,
    neuron_predicted = dfp.wide$neuron.bisque.scale,
    sample.id = rownames(dfp.wide),
    condition.id = rep("scale", nrow(dfp.wide)),
    replicate.id = rep("bisque", nrow(dfp.wide))
  ))
proportionsBulk <- rbind(proportionsBulk,
                         data.frame(
    glial_known = 1-dfp.wide$neuron.true,
    neuron_known = dfp.wide$neuron.true,
    glial__predicted = 1-dfp.wide$neuron.bisque.noscale,
    neuron_predicted = dfp.wide$neuron.bisque.noscale,
    sample.id = rownames(dfp.wide),
    condition.id = rep("no", nrow(dfp.wide)),
    replicate.id = rep("bisque", nrow(dfp.wide))
  ))

```

Cohort 1 bulk crossvalidate

```{r, eval = FALSE}
setwd("..")
setwd("..")
setwd("..")
load("./huuki-meyers-et-al-2023_dlpfc-cohort1/env/05_bulk/02_crossvalidate_script.RData")
proportionsBulkcv <- df.k2
rm(df.k2)
```

```{r, eval = FALSE}
# parse prep conditions
pBulkcv <- proportionsBulkcv
# format cell labels
colnames(pBulkcv)[c(1,2,12,13)] <- 
  c("glial_predicted", "neuron_predicted", 
    "glial_known", "neuron_known")
# format bulk id
pBulkcv$bulk.sample.id <- 
  ifelse(grepl("Nuc1", pBulkcv$bulk.sample.id), gsub("Nuc1", "Nuc", pBulkcv$bulk.sample.id),
         ifelse(grepl("Cyto1", pBulkcv$bulk.sample.id), gsub("Cyto1", "Cyto", pBulkcv$bulk.sample.id),
                ifelse(grepl("Bulk1", pBulkcv$bulk.sample.id), gsub("Bulk1", "Bulk", pBulkcv$bulk.sample.id), pBulkcv$bulk.sample.id)))

pBulkcv$cell.compartment <- 
  gsub(".*_", "", pBulkcv$bulk.sample.id)
# append prep condition
pBulkcvFilt <- pBulkcv[!pBulkcv$bulk.sample.condition=="NA",]
pBulkcv$bulk.sample.condition <-
  sapply(pBulkcv$bulk.sample.id, function(id){
    pBulkcvFilt[id,]$bulk.sample.condition})

pBulkcv$Librarycondition <- "NA"
pBulkcv$Librarycondition[
  grepl("polyA", pBulkcv$bulk.sample.condition)] <- "polyA"
pBulkcv$Librarycondition[
  grepl("RiboZero", pBulkcv$bulk.sample.condition)] <- "RiboZero"

# append cell compartment
pBulkcv$Cellcompartment <-
  ifelse(
    grepl("Nuc", pBulkcv$bulk.sample.condition), "Nuc", 
         ifelse(
           grepl("Cyto", pBulkcv$bulk.sample.condition), "Cyto", ifelse(
             grepl("Bulk", pBulkcv$bulk.sample.condition), "Bulk", "NA")))
pBulkcv$condition.id <- ifelse(pBulkcv$s.set.label=="s.null", "noscale", "scale")
# to numeric
pBulkcv$glial_known <- as.numeric(pBulkcv$glial_known)

# parse iterations
filterCV <- pBulkcv$crossvalidation=="validation"
filterCV <- filterCV & !pBulkcv$bulk.sample.condition=="NA"
filterCV <- filterCV & !pBulkcv$bulk.scale.type=="logcounts"
pBulkcvUnfilt <- pBulkcv
pBulkcv <- pBulkcv[filterCV,]
pBulkcv1 <- pBulkcv2 <- pBulkcv3 <- pBulkcv4 <- pBulkcv

# parse iterations
varExperiment <- pBulkcv1$bulk.sample.condition

# parse pBulkcv1
pBulkcv1$sample.id <- pBulkcv1$bulk.sample.condition
pBulkcv1$replicate.id <- pBulkcv1$experiment.type
pBulkcv1 <- pBulkcv1[,c(1,2,12,13,5,20,21)]
# parse pBulkcv2
pBulkcv2$sample.id <- pBulkcv2$Librarycondition
pBulkcv2$replicate.id <- pBulkcv2$experiment.type
pBulkcv2 <- pBulkcv2[,c(1,2,12,13,5,20,21)]
# parse pBulkcv3
pBulkcv3$sample.id <- pBulkcv3$Cellcompartment
pBulkcv3$replicate.id <- pBulkcv3$experiment.type
pBulkcv3 <- pBulkcv3[,c(1,2,12,13,5,20,21)]
```

Cohort 3 pseudobulk

```{r}
setwd("..")
setwd("..")
setwd("..")
load(file.path("./monaco-et-al-2019_pbmc-cohort1/env/01_pseudobulk/04_simulation.RData"))
proportionsCohort3 <- dfPseudobulk
# 
colnames(proportionsCohort3) <- c(
  "nonplasmablasts_predicted", "plasmablasts_predicted",
  "condition.id", "plasmablasts_known", "nonplasmablasts_known",
  "sample.id"
)
proportionsCohort3 <- proportionsCohort3[,c(1,2,4,5,6,3)]

```

Cohort 1, proportions shuffle

```{r}
setwd("..")
setwd("..")
setwd("..")
load(file.path("./huuki-meyers-et-al-2023_dlpfc-cohort1/env/03_shuffle/02_figde_script.RData"))

proportionsShuffle <- dfp.tall
colnames(proportionsShuffle) <- c(
  "glial_known", "neuron_known", "glial_predicted", "neuron_predicted",
  "sample.id.shuffle.pred", "condition.id", "sample.id", "matched.id"
)
proportionsShuffle <- proportionsShuffle[,c(1,2,3,4,7,6)]
rm(dfp.tall)
```

# Run RMSE calculations

K2 calculations

```{r} 
# cohorts 1 and 2,
listProportionsTablesK2 <- list(
  proportionsCohort1K2, proportionsCohort2K2, proportionsShuffle
)

# 2 cell types
experimentLabelsVectorK2 <- c("cohort1_k2", "cohort2_k2", "cohort1_shuffle")
cellTypesVectorK2 <- c("glial", "neuron")
rmseSuppTableK2Twotypes <- do.call(rbind, 
                   lapply(seq(length(listProportionsTablesK2)),
                                function(index){
  message(index)
  proportionsData <- listProportionsTablesK2[[index]]
  if(!"replicate.id" %in% colnames(proportionsData)){
    proportionsData$replicate.id <- "rep1"
  }
  dataLabel <- experimentLabelsVectorK2[[index]]
  returnList <- applyRmse(
    proportionsData, 
    cellTypesVector=cellTypesVectorK2
  )
  returnTable <- returnList$returnRmseTall
  returnTable$experimentLabel <- experimentLabelsVectorK2[index]
  return(returnTable)
})) |> 
  as.data.frame()

# 1 cell types
experimentLabelsVectorK2 <- c("cohort1_neuron_k2", "cohort2_neuron_k2", "cohort1_neuron_shuffle")
cellTypesVectorK2 <- c("neuron")
rmseSuppTableK2Neuron <- do.call(rbind, 
                   lapply(seq(length(listProportionsTablesK2)),
                                function(index){
  message(index)
  proportionsData <- listProportionsTablesK2[[index]]
  if(!"replicate.id" %in% colnames(proportionsData)){
    proportionsData$replicate.id <- "rep1"
  }
  dataLabel <- experimentLabelsVectorK2[[index]]
  returnList <- applyRmse(
    proportionsData, 
    cellTypesVector=c("neuron")
  )
  returnTable <- returnList$returnRmseTall
  returnTable$experimentLabel <- experimentLabelsVectorK2[index]
  return(returnTable)
})) |> 
  as.data.frame()
experimentLabelsVectorK2 <- c("cohort1_glial_k2", "cohort2_glial_k2", "cohort1_glial_shuffle")
cellTypesVectorK2 <- c("glial")
rmseSuppTableK2Glial <- do.call(rbind, 
                   lapply(seq(length(listProportionsTablesK2)),
                                function(index){
  message(index)
  proportionsData <- listProportionsTablesK2[[index]]
  if(!"replicate.id" %in% colnames(proportionsData)){
    proportionsData$replicate.id <- "rep1"
  }
  dataLabel <- experimentLabelsVectorK2[[index]]
  returnList <- applyRmse(
    proportionsData, 
    cellTypesVector=c("glial")
  )
  returnTable <- returnList$returnRmseTall
  returnTable$experimentLabel <- experimentLabelsVectorK2[index]
  return(returnTable)
})) |> 
  as.data.frame()

# cohort 3, 
# 2 cell types
cellTypesVectorK2 <- c("nonplasmablasts", "plasmablasts")
experimentLabelsVectorK2 <- c("cohort3_k2")
listProportionsTablesK2 <- list(
  proportionsCohort3
)
rmseSuppTableK2Cohort3Twotypes <- do.call(rbind, 
                   lapply(seq(length(listProportionsTablesK2)),
                                function(index){
  message(index)
  proportionsData <- listProportionsTablesK2[[index]]
  if(!"replicate.id" %in% colnames(proportionsData)){
    proportionsData$replicate.id <- "rep1"
  }
  dataLabel <- experimentLabelsVectorK2[[index]]
  returnList <- applyRmse(
    proportionsData, 
    cellTypesVector=cellTypesVectorK2
  )
  returnTable <- returnList$returnRmseTall
  returnTable$experimentLabel <- experimentLabelsVectorK2[index]
  return(returnTable)
})) |> 
  as.data.frame()

# 1 cell types
cellTypesVectorK2 <- c("nonplasmablasts")
experimentLabelsVectorK2 <- c("cohort3_nonplasmablasts_k2")
listProportionsTablesK2 <- list(
  proportionsCohort3
)
rmseSuppTableK2Cohort3Nonplasmablasts <- do.call(rbind, 
                   lapply(seq(length(listProportionsTablesK2)),
                                function(index){
  message(index)
  proportionsData <- listProportionsTablesK2[[index]]
  if(!"replicate.id" %in% colnames(proportionsData)){
    proportionsData$replicate.id <- "rep1"
  }
  dataLabel <- experimentLabelsVectorK2[[index]]
  returnList <- applyRmse(
    proportionsData, 
    cellTypesVector=cellTypesVectorK2
  )
  returnTable <- returnList$returnRmseTall
  returnTable$experimentLabel <- experimentLabelsVectorK2[index]
  return(returnTable)
})) |> 
  as.data.frame()

cellTypesVectorK2 <- c("plasmablasts")
experimentLabelsVectorK2 <- c("cohort3_plasmablasts_k2")
listProportionsTablesK2 <- list(
  proportionsCohort3
)
rmseSuppTableK2Cohort3Plasmablasts <- do.call(rbind, 
                   lapply(seq(length(listProportionsTablesK2)),
                                function(index){
  message(index)
  proportionsData <- listProportionsTablesK2[[index]]
  if(!"replicate.id" %in% colnames(proportionsData)){
    proportionsData$replicate.id <- "rep1"
  }
  dataLabel <- experimentLabelsVectorK2[[index]]
  returnList <- applyRmse(
    proportionsData, 
    cellTypesVector=cellTypesVectorK2
  )
  returnTable <- returnList$returnRmseTall
  returnTable$experimentLabel <- experimentLabelsVectorK2[index]
  return(returnTable)
})) |> 
  as.data.frame()

# try k2 from bulk
cellTypesVectorK2 <- c("glial", "neuron")
experimentLabelsVectorK2 <- c(
  "cohort1_k2_bulk_nnls", "cohort1_k2_bulk_music", "cohort1_k2_bulk_bisque")
listProportionsTablesK2 <- list(
  proportionsBulk[proportionsBulk$replicate.id=="nnls",],
  proportionsBulk[proportionsBulk$replicate.id=="music",],
  proportionsBulk[proportionsBulk$replicate.id=="bisque",]
)
rmseSuppTableK2Bulk <- do.call(rbind, 
                   lapply(seq(length(listProportionsTablesK2)),
                                function(index){
  message(index)
  proportionsData <- listProportionsTablesK2[[index]]
  if(!"replicate.id" %in% colnames(proportionsData)){
    proportionsData$replicate.id <- "rep1"
  }
  dataLabel <- experimentLabelsVectorK2[[index]]
  returnList <- applyRmse(
    proportionsData, 
    cellTypesVector=cellTypesVectorK2
  )
  returnTable <- returnList$returnRmseTall
  returnTable$experimentLabel <- experimentLabelsVectorK2[index]
  return(returnTable)
})) |> 
  as.data.frame()

# k2 from bulk cv
#cellTypesVectorK2 <- c("glial", "neuron")
#experimentLabelsVectorK2 <- c("cohort1_k2_bulkcv_bycondition",
#                              "cohort1_k2_bulkcv_bylibrary",
#                              "cohort1_k2_bulkcv_bycompartment")
#listProportionsTablesK2 <- list(
#  pBulkcv1, pBulkcv2, pBulkcv3
#)
#rmseSuppTableK2Bulkcv <- do.call(rbind, 
#                   lapply(seq(length(listProportionsTablesK2)),
#                                function(index){
#  message(index)
#  proportionsData <- listProportionsTablesK2[[index]]
#  if(!"replicate.id" %in% colnames(proportionsData)){
#    proportionsData$replicate.id <- "rep1"
#  }
#  dataLabel <- experimentLabelsVectorK2[[index]]
#  returnList <- applyRmse(
#    proportionsData, 
#    cellTypesVector=cellTypesVectorK2
#  )
#  returnTable <- returnList$returnRmseTall
#  returnTable$experimentLabel <- experimentLabelsVectorK2[index]
#  return(returnTable)
#})) |> 
#  as.data.frame()


```

K3 calculations

```{r}
# cohort 1 and cohort2

# 3 cell types
cellTypesVectorK3 <- c("glial", "Inhib", "Excit")
experimentLabelsVectorK3 <- c("cohort1_k3", "cohort2_k3")
listProportionsTablesK3 <- list(
  proportionsCohort1K3, proportionsCohort2K3
)
rmseSuppTableK3Threetypes <- do.call(rbind, 
                   lapply(seq(length(listProportionsTablesK3)),
                                function(index){
  message(index)
  proportionsData <- listProportionsTablesK3[[index]]
  if(!"replicate.id" %in% colnames(proportionsData)){
    proportionsData$replicate.id <- "rep1"
  }
  dataLabel <- experimentLabelsVectorK3[[index]]
  returnList <- applyRmse(
    proportionsData, 
    cellTypesVector=cellTypesVectorK3
  )
  returnTable <- returnList$returnRmseTall
  returnTable$experimentLabel <- experimentLabelsVectorK3[index]
  return(returnTable)
})) |> 
  as.data.frame()

# 1 cell types
cellTypesVectorK3 <- c("glial")
experimentLabelsVectorK3 <- c("cohort1_glial_k3", "cohort2_glial_k3")
rmseSuppTableK3Glial <- do.call(rbind, 
                   lapply(seq(length(listProportionsTablesK3)),
                                function(index){
  message(index)
  proportionsData <- listProportionsTablesK3[[index]]
  if(!"replicate.id" %in% colnames(proportionsData)){
    proportionsData$replicate.id <- "rep1"
  }
  dataLabel <- experimentLabelsVectorK3[[index]]
  returnList <- applyRmse(
    proportionsData, 
    cellTypesVector=cellTypesVectorK3
  )
  returnTable <- returnList$returnRmseTall
  returnTable$experimentLabel <- experimentLabelsVectorK3[index]
  return(returnTable)
})) |> 
  as.data.frame()

cellTypesVectorK3 <- c("Inhib")
experimentLabelsVectorK3 <- c("cohort1_Inhib_k3", "cohort2_Inhib_k3")
rmseSuppTableK3Inhib <- do.call(rbind, 
                   lapply(seq(length(listProportionsTablesK3)),
                                function(index){
  message(index)
  proportionsData <- listProportionsTablesK3[[index]]
  if(!"replicate.id" %in% colnames(proportionsData)){
    proportionsData$replicate.id <- "rep1"
  }
  dataLabel <- experimentLabelsVectorK3[[index]]
  returnList <- applyRmse(
    proportionsData, 
    cellTypesVector=cellTypesVectorK3
  )
  returnTable <- returnList$returnRmseTall
  returnTable$experimentLabel <- experimentLabelsVectorK3[index]
  return(returnTable)
})) |> 
  as.data.frame()

cellTypesVectorK3 <- c("Excit")
experimentLabelsVectorK3 <- c("cohort1_Excit_k3", "cohort2_Excit_k3")
rmseSuppTableK3Excit <- do.call(rbind, 
                   lapply(seq(length(listProportionsTablesK3)),
                                function(index){
  message(index)
  proportionsData <- listProportionsTablesK3[[index]]
  if(!"replicate.id" %in% colnames(proportionsData)){
    proportionsData$replicate.id <- "rep1"
  }
  dataLabel <- experimentLabelsVectorK3[[index]]
  returnList <- applyRmse(
    proportionsData, 
    cellTypesVector=cellTypesVectorK3
  )
  returnTable <- returnList$returnRmseTall
  returnTable$experimentLabel <- experimentLabelsVectorK3[index]
  return(returnTable)
})) |> 
  as.data.frame()
```

# Get aggregated table

Format values across tables on aggregation.

```{r}

roundDigits <- 3

listRmseTables <- list(
  rmseSuppTableK2Twotypes,
  rmseSuppTableK2Neuron,
  rmseSuppTableK2Glial,
  rmseSuppTableK3Threetypes,
  rmseSuppTableK3Excit,
  rmseSuppTableK3Inhib,
  rmseSuppTableK3Glial,
  rmseSuppTableK2Bulk,
  #rmseSuppTableK2Bulkcv,
  rmseSuppTableK2Cohort3Twotypes,
  rmseSuppTableK2Cohort3Nonplasmablasts,
  rmseSuppTableK2Cohort3Plasmablasts
)

# format rmse values
listRmseTables <- lapply(
  listRmseTables, function(tableIteration){
    tableIteration$rmseResult <- format(
      tableIteration$rmseResult, digits = roundDigits,
      scientific = TRUE
    )
    return(tableIteration)
  }
)

# get final table
rmseSuppTableOut <- do.call(rbind, 
                            lapply(listRmseTables, function(item){item})) |> as.data.frame()

```

# Save final RMSE table

Save formatted table for supplement

```{r}
rmseSupplement <- rmseSuppTableOut
# append columns
# append cohort
rmseSupplement$cohort <- ifelse(
  grepl("cohort1", rmseSupplement$experimentLabel), "Huuki-Myers et al 2023",
  ifelse(grepl("cohort2", rmseSupplement$experimentLabel), "Tran et al 2021",
         ifelse(grepl("cohort3", rmseSupplement$experimentLabel), "Monaco et al 2019", "NA")))
# append k
rmseSupplement$k <- ifelse(grepl("k2|shuffle", rmseSupplement$experimentLabel), "2",
                           ifelse(grepl("k3", rmseSupplement$experimentLabel), "3", "NA"))
# append num cell types
rmseSupplement$numCellTypes <- gsub("k", "", rmseSupplement$k)
conditionCellTypes <- grepl("neuron|glial|plasmablasts|nonplasmablasts|Inhib|Excit", rmseSupplement$condition)
rmseSupplement$numCellTypes[conditionCellTypes] <- 1

# append cell type strings
rmseSupplement$cellTypes <- ""
# k2
filterK2Brain <- rmseSupplement$k == "2" & grepl("cohort1|cohort2", rmseSupplement$experimentLabel)
filterK2Brain2Types <- filterK2Brain & rmseSupplement$numCellTypes=="2"
rmseSupplement$cellTypes[filterK2Brain2Types] <- "neuron;glial"
filterK2BrainNeuron <- filterK2Brain & grepl("neuron", rmseSupplement$experimentLabel)
rmseSupplement$cellTypes[filterK2BrainNeuron] <- "neuron"
filterK2BrainGlial <- filterK2Brain & grepl("glial", rmseSupplement$experimentLabel)
rmseSupplement$cellTypes[filterK2BrainGlial] <- "glial"
# k3
filterK3Brain <- rmseSupplement$k == "3" & grepl("cohort1|cohort2", rmseSupplement$experimentLabel)
filterK3Brain3Types <- filterK3Brain & rmseSupplement$numCellTypes==3
rmseSupplement$cellTypes[filterK3Brain3Types] <- "Excit;Inhib;glial"
filterK3BrainExcit <- filterK3Brain & grepl("Excit", rmseSupplement$experimentLabel)
rmseSupplement$cellTypes[filterK3BrainExcit] <- "Excit"
filterK3BrainInhib <- filterK3Brain & grepl("Inhib", rmseSupplement$experimentLabel)
rmseSupplement$cellTypes[filterK3BrainInhib] <- "Inhib"
filterK3BrainGlial <- filterK3Brain & grepl("glial", rmseSupplement$experimentLabel)
rmseSupplement$cellTypes[filterK3BrainGlial] <- "glial"
# k2 pbmc
filterK2Blood <- grepl("cohort3", rmseSupplement$experimentLabel)
filterK2Blood2Types <- filterK2Blood & rmseSupplement$numCellTypes==2
rmseSupplement$cellTypes[filterK2Blood2Types] <- "plasmablasts;nonplasmablasts"
filterK2BloodPlasmablasts <- filterK2Blood & grepl("plasmablasts", rmseSupplement$experimentLabel)
rmseSupplement$cellTypes[filterK2BloodPlasmablasts] <- "plasmablasts"
filterK2BloodNonplasmablasts <- filterK2Blood & grepl("nonplasmablasts", rmseSupplement$experimentLabel)
rmseSupplement$cellTypes[filterK2BloodNonplasmablasts] <- "nonplasmablasts"


filterK3Brain <- rmseSupplement$k == "k3" & grepl("cohort1|cohort2", rmseSupplement$experimentLabel)
rmseSupplement$cellTypes[filterK3Brain] <- "Excit;Inhib;glial"

# remove conditions
dim(rmseSupplement)
filterCondition <- grepl("sample.id", rmseSupplement$condition)
filterCondition <- filterCondition & grepl("cohort3", rmseSupplement$experimentLabel)
rmseSupplement <- rmseSupplement[!filterCondition,]
dim(rmseSupplement)
```

Write to file.

```{r}
setwd("..")
setwd("..")

# no round
write.csv(
  rmseSupplement, 
  file = "./outputs/02_aggregateRMSE/supp_table_formatted_rmse_write.csv",
  row.names = FALSE)

# with round
rmseSuppTableOutRound <- rmseSupplement
rmseSuppTableOutRound$rmseResult <- format(rmseSuppTableOutRound$rmseResult, digits = 3)
head(rmseSuppTableOutRound)
write.csv(
  rmseSuppTableOutRound, 
  file = "./outputs/02_aggregateRMSE/supp_table_formatted_rmse_round_write.csv",
  row.names = FALSE)

options(scipen=10)
write.csv(
  rmseSuppTableOutRound, 
  file = "./outputs/02_aggregateRMSE/supp_table_formatted_rmse_round_write_scipen.csv",
  row.names = FALSE)
options(scipen=0)  # restore the default
```

# Session info

```{r}
sessionInfo()
```
