---
title: "02_cell_proportions_bysample_s13"
author: "Sean Maden"
date: "2023-10-23"
output:
  pdf_document: default
  html_document: default
---

This notebook compares cell type labels and amounts between the 2 refernces 
(flow cytometry and mRNA yield), and available cell type labels from the 
ABIS-seq reference.

```{r setup, include=FALSE}
libv <- c("ggplot2", "dplyr", "reshape2")
sapply(libv, library, character.only = TRUE)
knitr::opts_chunk$set(echo = TRUE)
setwd("..")
setwd("..")
load("./env/02_abisseq/02_proportions_s13_script.Rdata")
```

# Label overlap

## Check cell label overlap

Cell labels, flow cytometry proportions, FC proportions

```{r}
labels.fc <- colnames(fc.proportions)
labels.fc
```
Cell labels, reference Zref

```{r}
labels.ref <- colnames(zref)
labels.ref
```

Intersect, FC proportions, Zref cell type labels

```{r}
labels.intersect <- intersect(labels.fc, labels.ref)
labels.intersect
```

```{r}
length(labels.intersect)
```

# Summaries -- flow cytometry proportions

## Sample labels

Unique sample lables

```{r}
unique.samples.fc <- unique(fc.proportions$Sample.Name)
unique.samples.fc
```

Number unique sample labels

```{r}
length(unique.samples.fc)
```

## Evaluating one of the S13 cohort samples (id: 453W)

```{r}
sample.index <- 1 # "453W"

row.data <- fc.proportions[sample.index, seq(2,ncol(fc.proportions))]
numeric.row.data <- as.numeric(row.data)
names(numeric.row.data) <- names(row.data)
```

Flow cytometry proportions summary

```{r}
summary(numeric.row.data)
```

Most abundant cell type label

```{r}
names(numeric.row.data[numeric.row.data==max(numeric.row.data)])
```

Least abundant cell type label

```{r}
names(numeric.row.data[numeric.row.data==min(numeric.row.data)])
```

Total of proportions

```{r}
sum(numeric.row.data)
```

### Evaluating proportions from total cell profile

Barplot of proportions, original scale

```{r}
df.plot <- data.frame(cell.type = names(numeric.row.data),
                      proportion = numeric.row.data)
df.plot$cell.type <- factor(df.plot$cell.type, 
                            levels = 
                              df.plot$cell.type[
                                order(df.plot$proportion)])

ggplot(df.plot, aes(y = cell.type, x = proportion)) + geom_bar(stat = "identity")

```

Barplot of proportions, scaled as fraction

```{r}
df.plot$proportion.rescale <- df.plot$proportion/sum(df.plot$proportion)

ggplot(df.plot, aes(y = cell.type, x = proportion.rescale)) + geom_bar(stat = "identity")

```

Barplots with color on label intersect

```{r}
df.plot$labels.intersect <- ifelse(
  df.plot$cell.type %in% labels.intersect, "TRUE", "FALSE")

ggplot(df.plot, 
       aes(y = cell.type, x = proportion.rescale, fill = labels.intersect)) + 
  geom_bar(stat = "identity")
```

### Evaluating proportions from reference cell profile

```{r}
df.plot <- df.plot[df.plot$labels.intersect=="TRUE",]
df.plot
```

```{r}
df.plot$proportion.rescale <- df.plot$proportion/sum(df.plot$proportion)

ggplot(df.plot, aes(y = cell.type, x = proportion.rescale)) + 
  geom_bar(stat = "identity")
```

## Evaluating all S13 samples on overlapping cell types

```{r}
filter.fc <- colnames(fc.proportions)[seq(2,ncol(fc.proportions))] %in%
  labels.intersect

df.plot.wide <- cbind(fc.proportions[,1,drop=F], fc.proportions[,filter.fc]) %>% as.data.frame()
#df.plot.tall <- t(df.plot.wide[,seq(2,ncol(df.plot.wide))]) %>% as.data.frame()
#colnames(df.plot.tall) <- df.plot.wide[,1]
#df.plot.tall$cell.type <- rownames(df.plot.tall)
df.plot.tall <- melt(df.plot.wide, id.vars = c("Sample.Name"))
colnames(df.plot.tall) <- c("sample.id", "cell.type", "proportion")
```

Cell proportions across S13 samples

```{r}
ggplot(df.plot.tall, aes(x = cell.type, y = proportion)) + 
  geom_jitter() + geom_boxplot(alpha = 0, color = "cyan") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Summaries -- mRNA yield

## Sample labels

Unique sample labels

```{r}
unique.samples.mrna <- unique(mrna.yield$Sample_Name)
unique.samples.mrna
```

Reformat labels

```{r}
unique.samples.mrna.format <- gsub("_.*", "", unique.samples.mrna)
unique.samples.mrna.format
```

Number unique sample labels

```{r}
length(unique.samples.mrna)
```

Intersect with FC proportions

```{r}
intersect.fc.mrna <- intersect(
  fc.proportions$Sample.Name, unique.samples.mrna.format)
intersect.fc.mrna
```

Number intersecting sample labels

```{r}
length(intersect.fc.mrna)
```

Outersecting labels FC

```{r}
outersect.fc <- fc.proportions$Sample.Name[
  !fc.proportions$Sample.Name %in% unique.samples.mrna.format]
outersect.fc
```

Outersecting labels mRNA yield

```{r}
outersect.mrna <- unique.samples.mrna.format[
  !unique.samples.mrna.format %in% fc.proportions$Sample.Name]
outersect.mrna
```

Variable summaries for column "yield_ng"

```{r}
summary(mrna.yield[,"yield_ng"])
```

Variable summaries for column "pg_cell"

```{r}
summary(mrna.yield[,"pg_cell"])
```

## Evaluating one of the S13 cohort samples (id: 453W)

```{r}
mrna.sample <- mrna.yield[mrna.yield$Sample_Name=="453W_PBMC",,drop=F]
```

Values

```{r}
mrna.sample
```

## mRNA yield across samples

```{r}
df.plot <- mrna.yield

ggplot(df.plot, aes(x = Sample_Name, y = yield_ng)) + geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Yield NG versus PG, original values

```{r}
ggplot(df.plot, aes(x = pg_cell, y = yield_ng)) + 
  geom_point(size = 4, alpha = 0.5) + geom_abline(slope = 1, intercept = 0)
```

Yield NG versus PG, ranks

```{r}
df.plot <- df.plot[order(df.plot$pg_cell),]
df.plot$pg_cell.rank <- seq(nrow(df.plot))
  
df.plot <- df.plot[order(df.plot$yield_ng),]
df.plot$yield_ng.rank <- seq(nrow(df.plot))

ggplot(df.plot, aes(x = pg_cell.rank, y = yield_ng.rank)) + 
  geom_point(size = 4, alpha = 0.5) + geom_abline(slope = 1, intercept = 0)
```


# Session info

```{r}
sessionInfo()
```
