---
title: "pseudobulk_experiment_stransform"
author: "Sean Maden"
date: '2022-10-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source.dpath <- file.path("deconvo_method-paper", "source")
source.fnv <- c("pb_methods.R")
for(fni in source.fnv){source(file.path(source.dpath, fni))}
```

# Pseudobulk example

To run a minimal pseudobulk experiment, we need example data consisting of the 
cell-level reference expression data, cell labels, cell ratio vectors, and scale
range. Many of these have default values, and they can be varied for experiments.

```{r}
```

# S-transformation experiment

We wish to evaluate the impact of performing transformations on the signature matrix data.
To do this, we need the transformation values and an outline of the experiment design.

Here, we focus on the cell type labels in the `celltypes` variable of `sef`, a 
SummarizedExperiment object. We can summarize the cell abundances by their labels
as follows.

```{r}
```


