---
title: "03_cell_size_summaries"
author: "Sean Maden"
date: "2023-11-22"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
libv <- c("lute", "ggplot2", "gridExtra", "reshape2", "ggforce")
sapply(libv, library, character.only = TRUE)
setwd("..")
setwd("..")
load("./env/01_pseudobulk/01_read_script.RData")
knitr::opts_chunk$set(echo = FALSE)
```

# Background

## Cell size calculations

Cell size scale factors obtained as the library sizes (column sums) from the 
RNA-seq reference in Monaco et al 2019 (Table S5).

Binarized cell sizes (Plasmablasts, Non-plasmablasts) for Non-plasmablasts were 
taken as the median of cell sizes.

## Cell type label details

Definitions of cell type labels for predictions.

```{r}
df.def <- data.frame(
  cell.type = c("B.Naive",
                "Basophils.LD",
                "MAIT",
                "mDCs",
                "Monocytes.C",
                "Monocytes.NC.I",
                "Neutrophils.LD",
                "NK",
                "pDCs",
                "Plasmablasts",
                "T.CD4.Naive",
                "T.CD8.Memory",
                "T.CD8.Naive",
                "T.gd.non.Vd2",
                "T.gd.Vd2"),
  description = c("Naive B-cells",
                  "Low-density Basophils",
                  "Mucosal-associated invariant T cells (MAIT)",
                  "Myeloid Dendritic Cells",
                  "Monocytes, type C",
                  "Monocytes, types NC and I combined",
                  "Low-density Neutrophils",
                  "Natural Killer",
                  "Plasmacytoid Dendritic Cells",
                  "Plasmablasts",
                  "CD4+ Naive T-cell",
                  "CD8+ Memory T-cell",
                  "CD8+ Naive T-cell",
                  "T lymphocyte, gamma delta, non-Vdelta2",
                  "T lymphocyte, gamma delta, Vdelta2")
)
knitr::kable(df.def)
```

# Analyses

## Cell sizes for reference cell types

Histogram and density plots of cell sizes

```{r}
dfPlotCellSize <- data.frame(cellType = names(cellSizes), size = as.numeric(cellSizes))
dfPlotCellSize$cellType <- factor(dfPlotCellSize$cellType, levels = dfPlotCellSize$cellType[order(dfPlotCellSize$size)])
plot1 <- ggplot(dfPlotCellSize, aes(x = size)) + geom_histogram() + xlab("Cell size") + ylab("Cell type count")
plot2 <- ggplot(dfPlotCellSize, aes(x = size)) + geom_density() + xlab("Cell size") + ylab("Cell amount density")
grid.arrange(plot1, plot2, nrow=2)
```

Stacked histogram of cell sizes

```{r}
plot3 <- ggplot(dfPlotCellSize, aes(x = size, fill = cellType)) + geom_histogram() + xlab("Cell size") + 
  ylab("Cell type count") + guides(fill=guide_legend(ncol=2))
plot3
```

Barplots of cell sizes, with color by cell size.

```{r}
ggplot(dfPlotCellSize, aes(x = cellType, y = size, fill = cellType)) + 
  geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle=45,hjust=1)) + 
  guides(fill=guide_legend(ncol=2))
```

## Cell sizes for Non-plasmablast cells

Barplots of cell sizes, with color by cell size.

```{r}
dfPlotNotPlasmablast <- dfPlotCellSize[!dfPlotCellSize$cellType=="Plasmablasts",]
dfPlotNotPlasmablast$conditionLabel <- "Non-plasmablasts"
dfPlotNotPlasmablast$dendriticCellType <- 
  ifelse(dfPlotNotPlasmablast$cellType %in% c("pDCs", "mDCs"), TRUE, FALSE)

ggplot(dfPlotNotPlasmablast, aes(x = cellType, y = size, fill = cellType)) + 
  geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle=45,hjust=1)) + 
  guides(fill=guide_legend(ncol=2))
```

Jitter points and boxplot for Non-plasmablast cell sizes.

```{r, fig.width=8}
ggplot(
  dfPlotNotPlasmablast, aes(x = conditionLabel, y = size)) + 
  geom_boxplot(alpha = 0, color = "black") + 
  geom_jitter(aes(color = cellType, shape = dendriticCellType), size = 4)
```

## Cell sizes for Plasmablasts and Non-plasmablasts

```{r}
dfPlotTwoGroups <- dfPlotCellSize
dfPlotTwoGroups$condition <- 
  ifelse(dfPlotTwoGroups$cellType=="Plasmablasts", "Plasmablasts", "Non-plasmablasts")
dfPlotTwoGroups <- 
  aggregate(size ~ condition, data = dfPlotTwoGroups, FUN = median)
knitr::kable(dfPlotTwoGroups)
```

```{r}
plot1 <- 
  ggplot(dfPlotTwoGroups, aes(x = condition, y = size, fill = condition)) + 
  geom_bar(stat = "identity") + xlab("Cell type") + ylab("Cell size") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")
plot1
```
# Session info

```{r}
sessionInfo()
```