---
title: "01_read"
author: "Sean Maden"
date: "2023-10-16"
bibliograhy: bibliography.bib
output:
  pdf_document: default
  html_document: default
---

Summaries of expression tables read from GEO. Expression scale in transcripts 
per million (TPM). Data was published in @monaco_rna-seq_2019.

```{r setup, include=FALSE}
libv <- c("ggplot2", "dplyr")
sapply(libv, library, character.only = TRUE)
knitr::opts_chunk$set(echo = TRUE)
setwd("..")
setwd("..")
load("./env/01_tpm_summaries/01_read_script.RData")
```

# Summaries

# TPM

Summary statistics and plots. Uses the scale for the GEO record, or TPM.

Library sizes

```{r}
ggplot(cd, aes(x = tissue.type.detail, y = library.size, color = tissue.type)) + 
  geom_jitter() + geom_boxplot(alpha = 0, color = "cyan") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Library size")
```

Median expression

```{r}
ggplot(cd, aes(x = tissue.type.detail, y = median.expression, color = tissue.type)) + 
  geom_jitter() + geom_boxplot(alpha = 0, color = "cyan") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Median expression")
```

Mean expression

```{r}
ggplot(cd, aes(x = tissue.type.detail, y = mean.expression, color = tissue.type)) + 
  geom_jitter() + geom_boxplot(alpha = 0, color = "cyan") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Mean expression")
```

SD expression

```{r}
ggplot(cd, aes(x = tissue.type.detail, y = sd.expression, color = tissue.type)) + 
  geom_jitter() + geom_boxplot(alpha = 0, color = "cyan") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Mean expression")
```

Number of zero-expression genes

```{r}
ggplot(cd, aes(x = tissue.type.detail, y = num.zero.expression, color = tissue.type)) + 
  geom_jitter() + geom_boxplot(alpha = 0, color = "cyan") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  ggtitle("Number zero-expression genes")
```

# Log2(TPM+1)

Summary statistics and plots. Uses the scale for the study analysis, or Log2 TPM + 1.


Library sizes

```{r}
ggplot(cd, aes(x = tissue.type.detail, y = library.size, color = tissue.type)) + 
  geom_jitter() + geom_boxplot(alpha = 0, color = "cyan") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Library size")
```

Median expression

```{r}
ggplot(cd, aes(x = tissue.type.detail, y = median.expression, color = tissue.type)) + 
  geom_jitter() + geom_boxplot(alpha = 0, color = "cyan") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Median expression")
```

Mean expression

```{r}
ggplot(cd, aes(x = tissue.type.detail, y = mean.expression, color = tissue.type)) + 
  geom_jitter() + geom_boxplot(alpha = 0, color = "cyan") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Mean expression")
```

SD expression

```{r}
ggplot(cd, aes(x = tissue.type.detail, y = sd.expression, color = tissue.type)) + 
  geom_jitter() + geom_boxplot(alpha = 0, color = "cyan") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Mean expression")
```

Number of zero-expression genes

```{r}
ggplot(cd, aes(x = tissue.type.detail, y = num.zero.expression, color = tissue.type)) + 
  geom_jitter() + geom_boxplot(alpha = 0, color = "cyan") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  ggtitle("Number zero-expression genes")
```

# Session info

```{r}
sessionInfo()
```

# Works Cited
